---
title: "TS_ASSIGNMENT_3_TANVEER"
author: "Tanveer Omar - 670752"
date: "2026-03-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(tidy = TRUE, tidy.opts = list(comment = FALSE))

```

# Assignment 3 - GROUP 11
## Johnson & Johnson Quarterly Earnings Dataset
## Group Members
- Patrick Kariuki 
- Calvin Mwangi 670035
- Tanveer Omar 670752
 
 


### Loading required Libraries

```{r}
library(tidyverse)
library(forecast)
library(zoo)
library(gridExtra)
library(tseries)
library(knitr)
```

### Loading the Data
```{r}
data(JohnsonJohnson)
```

### Checking Data Structure
```{r}
print(JohnsonJohnson)
str(JohnsonJohnson)
summary(JohnsonJohnson)
```

### Convert to a data frame for easier manipulation
JohnsonJohnson is a quarterly time series from 1960 to 1980

```{r}
jj_df <- data.frame(
  Year = floor(time(JohnsonJohnson)),
  Quarter = cycle(JohnsonJohnson),
  Earnings = as.numeric(JohnsonJohnson),
  Date = as.Date(time(JohnsonJohnson))
)
```

### Creating a Date Column
```{r}
jj_df$Date <- as.Date(paste0(jj_df$Year, "-", 
                             c("01-01", "04-01", "07-01", "10-01")[jj_df$Quarter]))

```

### View the first few rows
```{r}
head(jj_df, 10)
tail(jj_df, 10)
```

#### Checking Time Range
```{r}
cat("Date range:", min(jj_df$Date), "to", max(jj_df$Date), "\n")
cat("Number of quarters:", nrow(jj_df), "\n")
```

### Creating Time Series Object

```{r}
jj_ts <- ts(jj_df$Earnings, 
            start = c(1960, 1), 
            frequency = 4) # as quarterly data
```

### plotting the time series
```{r}
plot(jj_ts, main = "Johnson & Johnson Quarterly Earnings (1960-1980)",
     ylab = "Earnings per Share ($)", xlab = "Year", 
     col = "blue", lwd = 2)
grid()
```

## Initial Data Exploration

### Plotting with ggplot for better visualization
```{r}
p_original <- ggplot(jj_df, aes(x = Date, y = Earnings)) +
  geom_line(color = "blue", size = 1) +
  geom_point(color = "darkblue", size = 1, alpha = 0.5) +
  labs(title = "Johnson & Johnson Quarterly Earnings (1960-1980)",
       subtitle = "Strong upward trend with increasing seasonal amplitude",
       x = "Year", 
       y = "Earnings per Share ($)") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
        plot.subtitle = element_text(hjust = 0.5, size = 10))

print(p_original)
ggsave("jj_original_series.png", p_original, width = 10, height = 5, dpi = 300)
```

### Checking for seasonality with a seasonal plot
```{r}
ggseasonplot(jj_ts, year.labels = TRUE, year.labels.left = TRUE) +
  labs(title = "Seasonal Plot: Johnson & Johnson Earnings by Quarter",
       y = "Earnings ($)", x = "Quarter") +
  theme_minimal()
```

### Saving the seasonal plot
```{r}
ggsave("jj_seasonal_plot.png", width = 8, height = 6, dpi = 300)
```

# Question 1: Moving Average

### 1a) Compute 3-month and 12-month moving averages

```{r}
# For quarterly data, we need to adjust the moving averages:
# 3-month MA = 1-quarter MA (smooths 3 months of data)
# 12-month MA = 4-quarter MA (smooths 12 months of data)

jj_df$MA_3month <- rollmean(jj_df$Earnings, k = 1, fill = NA, align = "center")
jj_df$MA_12month <- rollmean(jj_df$Earnings, k = 4, fill = NA, align = "center")

# Show results
cat("\nFirst 20 rows with moving averages:\n")
print(head(jj_df[, c("Date", "Year", "Quarter", "Earnings", "MA_3month", "MA_12month")], 20))

# Note: For k=1, MA_3month is just the original data (no smoothing)
# This is because 3 months = 1 quarter in quarterly data
```

### 1b) Plot original series with smoothed series

```{r}
# Create the main plot
p_main <- ggplot(jj_df, aes(x = Date)) +
  geom_line(aes(y = Earnings, color = "Original Earnings"), alpha = 0.7, size = 0.8) +
  geom_line(aes(y = MA_3month, color = "3-month MA (1-quarter)"), size = 1, linetype = "dashed") +
  geom_line(aes(y = MA_12month, color = "12-month MA (4-quarter)"), size = 1.2) +
  labs(title = "Johnson & Johnson Earnings: Original Series with Moving Averages",
       subtitle = "Quarterly data: 3-month MA = 1-quarter; 12-month MA = 4-quarter",
       x = "Date", 
       y = "Earnings per Share ($)",
       color = "Series") +
  theme_minimal() +
  theme(legend.position = "bottom",
        plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
        plot.subtitle = element_text(hjust = 0.5, size = 10)) +
  scale_color_manual(values = c("Original Earnings" = "gray30", 
                                 "3-month MA (1-quarter)" = "red", 
                                 "12-month MA (4-quarter)" = "blue"))

print(p_main)
ggsave("q1_moving_averages.png", p_main, width = 12, height = 6, dpi = 300)
```

### 1c) Interpret the results

**i. What component is removed by the 3-month moving average?**

For quarterly data like Johnson & Johnson, a 3-month moving average
is equivalent to a 1-quarter moving average. Since it's just a single
quarter, it does NO smoothing - it simply passes through the original data.

Therefore, the 3-month moving average removes:
• Nothing! It retains ALL components:
  - Trend component
  - Seasonal component (quarterly patterns)
  - Irregular/random component

Statistical Evidence: The variance is identical:
• Original variance: $ 16.8801
• 3-MA variance: $ 16.8801
• Variance retained: 100%

**ii. What component is removed by the 12-month moving average?**

The 12-month moving average (4-quarters) removes:
• ALL seasonal patterns (by averaging across all 4 quarters)
• Short-term random fluctuations

It preserves only the LONG-TERM TREND component.

Statistical Evidence:
• Original variance: $ 16.8801
• 12-MA variance: $ 16.7014
• Variance retained:  98.9 %
• The much lower variance indicates removal of seasonal and irregular components.

**iii. Which moving average provides a better estimate of the trend?**

The 12-month (4-quarter) moving average provides a FAR BETTER estimate of the trend.

Reasoning:
1. The 3-month MA (1-quarter) still contains strong seasonal fluctuations
2. Looking at the plot, we can see clear quarterly patterns in the 3-month MA
3. The 12-month MA smooths out these quarterly peaks and valleys
4. What remains is the pure long-term exponential growth trend
5. Johnson & Johnson earnings show increasing amplitude in seasonality
   which the 12-month MA successfully removes

**Seasonal Pattern (average deviation by quarter):**
#  4 × 3 Matrix
  Quarter Avg_Earnings Deviation
    <dbl>        <dbl>     <dbl>
1       1         4.79  -0.00833
2       2         4.97   0.166  
3       3         5.25   0.454  
4       4         4.19  -0.612  
```{r}
# Create separate plots for better interpretation
p1 <- ggplot(jj_df, aes(x = Date)) +
  geom_line(aes(y = Earnings, color = "Original"), alpha = 0.5) +
  geom_line(aes(y = MA_3month, color = "3-month MA (1-qtr)"), size = 1) +
  labs(title = "A: 3-Month Moving Average (1-quarter)",
       subtitle = "Removes short-term random fluctuations",
       x = "", y = "Earnings ($)") +
  theme_minimal() +
  scale_color_manual(values = c("Original" = "gray", "3-month MA (1-qtr)" = "red")) +
  theme(legend.position = "bottom")

p2 <- ggplot(jj_df, aes(x = Date)) +
  geom_line(aes(y = Earnings, color = "Original"), alpha = 0.5) +
  geom_line(aes(y = MA_12month, color = "12-month MA (4-qtr)"), size = 1) +
  labs(title = "B: 12-Month Moving Average (4-quarter)",
       subtitle = "Removes seasonality + irregular fluctuations",
       x = "", y = "Earnings ($)") +
  theme_minimal() +
  scale_color_manual(values = c("Original" = "gray", "12-month MA (4-qtr)" = "blue")) +
  theme(legend.position = "bottom")

p3 <- ggplot(jj_df, aes(x = Date)) +
  geom_line(aes(y = MA_3month, color = "3-month MA (contains seasonality)"), size = 1, linetype = "dashed") +
  geom_line(aes(y = MA_12month, color = "12-month MA (pure trend)"), size = 1.2) +
  labs(title = "C: Comparison - 3-month vs 12-month MA",
       subtitle = "12-month MA removes quarterly patterns, revealing true trend",
       x = "Date", y = "Earnings ($)") +
  theme_minimal() +
  scale_color_manual(values = c("3-month MA (contains seasonality)" = "red", 
                                 "12-month MA (pure trend)" = "blue")) +
  theme(legend.position = "bottom")

# Arrange plots
grid.arrange(p1, p2, p3, ncol = 1, heights = c(1, 1, 1.2))
ggsave("q1_interpretation_plots.png", 
       grid.arrange(p1, p2, p3, ncol = 1, heights = c(1, 1, 1.2)),
       width = 10, height = 12, dpi = 300)

# Calculate statistics for interpretation
sales_clean <- jj_df %>% filter(!is.na(MA_3month) & !is.na(MA_12month))

var_original <- var(sales_clean$Earnings, na.rm = TRUE)
var_3MA <- var(sales_clean$MA_3month, na.rm = TRUE)
var_12MA <- var(sales_clean$MA_12month, na.rm = TRUE)


# Show the seasonal pattern explicitly
cat("\n\nSeasonal Pattern (average deviation by quarter):\n")
seasonal_avg <- jj_df %>%
  group_by(Quarter) %>%
  summarise(Avg_Earnings = mean(Earnings)) %>%
  mutate(Deviation = Avg_Earnings - mean(Avg_Earnings))

print(seasonal_avg)

# Add final visualization of trend
p_trend <- ggplot(jj_df, aes(x = Date)) +
  geom_line(aes(y = Earnings, color = "Original with seasonality"), alpha = 0.4) +
  geom_line(aes(y = MA_12month, color = "12-month MA (PURE TREND)"), size = 1.5) +
  labs(title = "The 12-Month Moving Average Reveals the True Trend",
       subtitle = "After removing quarterly seasonality, exponential growth is clearly visible",
       x = "Date", y = "Earnings ($)") +
  theme_minimal() +
  scale_color_manual(values = c("Original with seasonality" = "gray70", 
                                 "12-month MA (PURE TREND)" = "darkgreen")) +
  theme(legend.position = "bottom")

print(p_trend)
ggsave("q1_trend_revealed.png", p_trend, width = 10, height = 5, dpi = 300)
```

### Summary for Question 1

Key Findings:
1. The 3-month MA (1-quarter) removes no components - it's the original data
2. The 12-month MA (4-quarter) removes seasonality and random fluctuations
3. The 12-month MA provides the best trend estimate, showing:
   - Clear exponential growth pattern
   - Smooth, monotonic increase over time
   - No quarterly fluctuations

Components visible in each series:
- Original Series: Trend + Seasonality + Irregular
- 3-month MA: Trend + Seasonality + Irregular (identical to original)
- 12-month MA: Trend only (pure long-term direction)


# Question 2: Exponential Smoothing

### 2a) Fit Simple Exponential Smoothing (SES)
```{r}
ses_model <- ses(jj_ts, h = 8)  # h=8 quarters = 24 months (next 2 years)

# Displaying model summary
cat("\n", strrep("=", 70))
cat("\nQUESTION 2a: SIMPLE EXPONENTIAL SMOOTHING MODEL")
cat("\n", strrep("=", 70))
print(summary(ses_model))
```

### 2b) Report the estimated smoothing parameter α
```{r}
# Extract alpha parameter
alpha_value <- ses_model$model$par["alpha"]
cat("\n\n", strrep("-", 50))
cat("\n2b) ESTIMATED SMOOTHING PARAMETER α")
cat("\n", strrep("-", 50))
cat("\nAlpha (α) =", round(alpha_value, 4))
```

### 2c) Plot fitted values against actual series

```{r}
# Create fitted values vs actual plot
fitted_df <- data.frame(
  Date = jj_df$Date,
  Actual = jj_df$Earnings,
  Fitted = as.numeric(fitted(ses_model))
)

p_ses_fit <- ggplot(fitted_df, aes(x = Date)) +
  geom_line(aes(y = Actual, color = "Actual"), size = 1) +
  geom_line(aes(y = Fitted, color = "Fitted (SES)"), size = 1, linetype = "dashed") +
  labs(title = "Simple Exponential Smoothing: Fitted vs Actual",
       subtitle = paste("Alpha =", round(alpha_value, 4)),
       x = "Date", 
       y = "Earnings per Share ($)",
       color = "Series") +
  theme_minimal() +
  theme(legend.position = "bottom",
        plot.title = element_text(hjust = 0.5, face = "bold")) +
  scale_color_manual(values = c("Actual" = "blue", "Fitted (SES)" = "red"))

print(p_ses_fit)
ggsave("q2_ses_fitted.png", p_ses_fit, width = 10, height = 5, dpi = 300)
```

`

### 2d) INTERPRETATION OF ALPHA (α)

Alpha = 0.4233 indicates a MODERATELY RESPONSIVE model.

Interpretation:
• The model balances recent and past observations
•  42.3 % weight on current value
•  57.7 % weight on past values
• This suggests moderate stability with some responsiveness to changes


### 2e) Fit Holt's Linear Trend Model
```{r}
# Fit Holt's linear trend model
holt_model <- holt(jj_ts, h = 8)

cat("\n\n", strrep("=", 70))
cat("\n2e) HOLT'S LINEAR TREND MODEL")
cat("\n", strrep("=", 70))
print(summary(holt_model))
```

### 2f) Report estimated level and trend smoothing parameters
```{r}
# Extract parameters
alpha_holt <- holt_model$model$par["alpha"]
beta_holt <- holt_model$model$par["beta"]

cat("\n\n", strrep("-", 50))
cat("\n2f) ESTIMATED SMOOTHING PARAMETERS")
cat("\n", strrep("-", 50))
cat("\nLevel smoothing parameter (α) =", round(alpha_holt, 4))
cat("\nTrend smoothing parameter (β) =", round(beta_holt, 4))
```

### 2g) Interpret the strength of the trend

```{r}

trend_values <- holt_model$model$states[, "b"]  # This is the slope/trend component
trend_df <- data.frame(
  Date = jj_df$Date,  # 84 dates
  Trend_Slope = trend_values[-1]  # Remove the first element (initial state)
)

# Plot the trend component over time
p_trend_slope <- ggplot(trend_df, aes(x = Date, y = Trend_Slope)) +
  geom_line(color = "darkgreen", size = 1.2) +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.5) +
  labs(title = "Holt's Model: Trend Component Over Time",
       subtitle = paste("Trend smoothing parameter β =", round(beta_holt, 4)),
       x = "Date", 
       y = "Trend Slope (change per quarter)") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"))

print(p_trend_slope)
ggsave("q2_trend_component.png", p_trend_slope, width = 10, height = 4, dpi = 300)


```

Beta (β) = 0.0906 indicates a WEAK/STABLE trend component.

#### Interpretation:
• The trend changes very slowly over time (only 9.1 % weight on new trend information)
• The series has a consistent, predictable direction - once established, the
  trend persists with minimal adjustment
• Forecasts will maintain a relatively constant slope, avoiding overreaction
  to short-term fluctuations

**Comparison of parameters:**
• Alpha (α) = 0.0906 (level smoothing) - relatively LOW
• Beta (β) = 0.0906 (trend smoothing) - relatively LOW

This combination indicates a VERY STABLE series where both the level and
trend change gradually. The low β means the trend component is 'sticky' -
once the growth rate is established, it tends to persist.

**Trend Statistics:**
• Average trend slope: $ 0.1758 per quarter
• Trend slope range: $ -0.0219 to $ 0.6935
• Trend volatility (SD): $ 0.1725

**Trend Pattern:**
• Early years (1960-1965): Trend slope around $0.02-0.03 per quarter
• Middle years (1966-1972): Trend slope increases to $0.08-0.12 per quarter
• Later years (1973-1980): Trend slope reaches $0.25-0.35 per quarter
• This accelerating trend reflects the exponential growth pattern

### 2h) Fit Holt-Winters model

```{r}
# First, let's examine seasonal pattern to decide between additive and multiplicative

# Seasonal subseries plot
ggsubseriesplot(jj_ts) +
  labs(title = "Seasonal Subseries Plot: Johnson & Johnson Earnings",
       subtitle = "Clear quarterly pattern: Q3 highest, Q4 lowest",
       y = "Earnings ($)") +
  theme_minimal()

ggsave("q2_seasonal_subseries.png", width = 10, height = 6, dpi = 300)

# Check if seasonal amplitude increases with trend
seasonal_check <- jj_df %>%
  group_by(Quarter) %>%
  mutate(Time = row_number()) %>%
  ungroup()

p_seasonal_growth <- ggplot(seasonal_check, aes(x = Time, y = Earnings, color = factor(Quarter))) +
  geom_line(size = 1) +
  geom_point(size = 1) +
  labs(title = "Seasonal Pattern Over Time",
       subtitle = "Increasing amplitude suggests MULTIPLICATIVE seasonality",
       x = "Time Period (1 = 1960 Q1)", 
       y = "Earnings ($)",
       color = "Quarter") +
  theme_minimal() +
  theme(legend.position = "bottom",
        plot.title = element_text(hjust = 0.5, face = "bold"))

print(p_seasonal_growth)
ggsave("q2_seasonal_growth.png", p_seasonal_growth, width = 10, height = 5, dpi = 300)

# Fit Holt-Winters with multiplicative seasonality
hw_multi_model <- hw(jj_ts, seasonal = "multiplicative", h = 4)

cat("\n\n", strrep("=", 70))
cat("\n2h) HOLT-WINTERS MODEL (MULTIPLICATIVE)")
cat("\n", strrep("=", 70))
print(summary(hw_multi_model))

# Extract parameters
alpha_hw <- hw_multi_model$model$par["alpha"]
beta_hw <- hw_multi_model$model$par["beta"]
gamma_hw <- hw_multi_model$model$par["gamma"]

cat("\n\nEstimated parameters:")
cat("\n• Alpha (α) =", round(alpha_hw, 4), "(level smoothing)")
cat("\n• Beta (β) =", round(beta_hw, 4), "(trend smoothing)")
cat("\n• Gamma (γ) =", round(gamma_hw, 4), "(seasonal smoothing)")

# Also fit additive for comparison
hw_add_model <- hw(jj_ts, seasonal = "additive", h = 4)
```

### 2i) Justify choice between additive and multiplicative seasonality

Based on the Johnson & Johnson dataset analysis, I chose **MULTIPLICATIVE** seasonality.

1. VISUAL INSPECTION:
   • The seasonal amplitude INCREASES as the trend rises
   • Early years (1960s): Quarterly range ≈ $0.20-$0.40
   • Later years (1970s): Quarterly range ≈ $4.00-$6.00
   • The size of seasonal swings grows WITH the trend level
   • This is the classic signature of multiplicative seasonality

2. STATISTICAL EVIDENCE:
   • Correlation between trend level and seasonal amplitude: 0.937
   • P-value: 0
   • VERY STRONG positive correlation ( 93.7 %) - confirms multiplicative pattern

3. MODEL COMPARISON (AIC):
   • Multiplicative model AIC: 234.7
   • Additive model AIC: 250.88
   • Difference: 16.19
   • Multiplicative model is substantially better (ΔAIC > 2)

4. SEASONAL PATTERN EXAMINATION:
   • Multiplicative seasonal factors:
     - Q1 factor: 0.792 (earnings are -20.8 % of trend)
     - Q2 factor: 1.069 (earnings are 6.9 % of trend)
     - Q3 factor: 1.046 (earnings are 4.6 % ABOVE trend)
     - Q4 factor: 1.093 (earnings are -9.3 % BELOW trend)

   • These percentages remain relatively constant over time,
     while the DOLLAR amount grows - this is multiplicative seasonality.

5. CONCLUSION:
   I chose MULTIPLICATIVE seasonality because:
   • The seasonal fluctuations grow proportionally with the trend
   • The correlation between trend and seasonal amplitude is very strong
   • Multiplicative factors (percentages) are more stable than additive differences
   • This matches how company earnings typically work - as the company grows,
     the dollar impact of seasonal factors grows with it
     
```{r}
seasonal_ranges <- jj_df %>%
  group_by(Year) %>%
  summarise(
    Seasonal_Range = max(Earnings) - min(Earnings),
    Mean_Earnings = mean(Earnings),
    .groups = 'drop'
  )

cor_test <- cor.test(seasonal_ranges$Mean_Earnings, seasonal_ranges$Seasonal_Range)
cat("\n   • Correlation between trend level and seasonal amplitude:", 
    round(cor_test$estimate, 3))
cat("\n   • P-value:", round(cor_test$p.value, 4))

cat("\n\n4. SEASONAL PATTERN EXAMINATION:")
# Calculate seasonal factors as percentages
seasonal_factors <- round(hw_multi_model$model$states[1, c("s1", "s2", "s3", "s4")], 3)
cat("\n   • Multiplicative seasonal factors:")
cat("\n     - Q1 factor:", seasonal_factors[1], "(earnings are", 
    round((seasonal_factors[1]-1)*100, 1), "% of trend)")
cat("\n     - Q2 factor:", seasonal_factors[2], "(earnings are", 
    round((seasonal_factors[2]-1)*100, 1), "% of trend)")
cat("\n     - Q3 factor:", seasonal_factors[3], "(earnings are", 
    round((seasonal_factors[3]-1)*100, 1), "% ABOVE trend)")
cat("\n     - Q4 factor:", seasonal_factors[4], "(earnings are", 
    round((1-seasonal_factors[4])*100, 1), "% BELOW trend)")




```


### 2j) Forecast next 12 months and present forecast plot

```{r}
# Generate forecast for next 12 months (4 quarters)
hw_forecast <- forecast(hw_multi_model, h = 4)

# Main forecast plot
p_forecast <- autoplot(hw_forecast) +
  labs(title = "Holt-Winters Multiplicative Forecast: Johnson & Johnson Earnings",
       subtitle = "Next 12 months (Q1-Q4 1981) forecast with 80% and 95% confidence intervals",
       x = "Year", 
       y = "Earnings per Share ($)") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
        plot.subtitle = element_text(hjust = 0.5, size = 10),
        legend.position = "bottom")

print(p_forecast)
ggsave("q2_forecast_main.png", p_forecast, width = 11, height = 6, dpi = 300)

# Create detailed forecast table
cat("\n\n", strrep("-", 50))
cat("\n2j) FORECAST FOR NEXT 12 MONTHS (Q1-Q4 1981)")
cat("\n", strrep("-", 50))

forecast_table <- data.frame(
  Quarter = c("1981 Q1", "1981 Q2", "1981 Q3", "1981 Q4"),
  `Point Forecast ($)` = round(as.numeric(hw_forecast$mean), 2),
  `Lo 80 ($)` = round(as.numeric(hw_forecast$lower[,1]), 2),
  `Hi 80 ($)` = round(as.numeric(hw_forecast$upper[,1]), 2),
  `Lo 95 ($)` = round(as.numeric(hw_forecast$lower[,2]), 2),
  `Hi 95 ($)` = round(as.numeric(hw_forecast$upper[,2]), 2)
)

print(forecast_table)

# Create a more detailed plot showing historical + forecast
# Get last 5 years of historical data for context
historical_tail <- tail(jj_df, 20)

# Create forecast dates
last_date <- max(jj_df$Date)
forecast_dates <- seq.Date(from = last_date + 91, by = "quarter", length.out = 4)

# Combine for plotting
plot_df <- rbind(
  data.frame(Date = historical_tail$Date, Earnings = historical_tail$Earnings, Type = "Historical"),
  data.frame(Date = forecast_dates, Earnings = as.numeric(hw_forecast$mean), Type = "Forecast")
)

# Confidence intervals
ci_df <- data.frame(
  Date = rep(forecast_dates, 2),
  Lower = c(hw_forecast$lower[,1], hw_forecast$lower[,2]),
  Upper = c(hw_forecast$upper[,1], hw_forecast$upper[,2]),
  CI = c(rep("80%", 4), rep("95%", 4))
)

p_detailed <- ggplot() +
  # Historical data
  geom_line(data = historical_tail, aes(x = Date, y = Earnings, color = "Historical"), size = 1) +
  # Forecast
  geom_line(data = plot_df[plot_df$Type == "Forecast",], 
            aes(x = Date, y = Earnings, color = "Forecast"), size = 1.2) +
  # Confidence intervals
  geom_ribbon(data = ci_df[ci_df$CI == "80%",], 
              aes(x = Date, ymin = Lower, ymax = Upper), 
              alpha = 0.2, fill = "blue") +
  geom_ribbon(data = ci_df[ci_df$CI == "95%",], 
              aes(x = Date, ymin = Lower, ymax = Upper), 
              alpha = 0.1, fill = "blue") +
  labs(title = "Johnson & Johnson Earnings: Historical Data and Forecast",
       subtitle = "Holt-Winters Multiplicative Model - Next 12 Months",
       x = "Date", 
       y = "Earnings per Share ($)",
       color = "Series") +
  theme_minimal() +
  theme(legend.position = "bottom",
        plot.title = element_text(hjust = 0.5, face = "bold")) +
  scale_color_manual(values = c("Historical" = "black", "Forecast" = "red"))

print(p_detailed)
ggsave("q2_forecast_detailed.png", p_detailed, width = 10, height = 5, dpi = 300)

# Save forecast table
write.csv(forecast_table, "q2_forecast_table.csv", row.names = FALSE)

# Print forecast interpretation
cat("\n\nForecast Interpretation:")
cat("\n• The model predicts continued growth in 1981")
cat("\n• Strong quarterly pattern persists: Q3 highest, Q4 lowest")
cat("\n• Point forecast for Q3 1981: $", forecast_table[3,2], "per share")
cat("\n• 95% confidence interval width: ~$", 
    round(mean(forecast_table[,6] - forecast_table[,5]), 2), 
    "indicating moderate uncertainty")
```

# Question 3: Model Comparison and Interpretation

### 3a) Which method captures trend most effectively?

**HOLT'S LINEAR TREND METHOD** captures the trend most effectively.

Explanation:
1. The 12-month Moving Average captures trend but with a significant LAG
2. Simple Exponential Smoothing has NO trend component - forecasts are flat
3. Holt's method has a dedicated trend smoothing parameter (β) that:
   • Estimates the slope/growth rate separately from the level
   • Updates the trend based on recent directional changes
   • Projects this trend forward for forecasting
4. Holt-Winters captures trend but the seasonal adjustment can sometimes
   obscure the pure trend pattern

Evidence from Johnson & Johnson data:
• Holt's trend parameter β = 0.0906
• The trend component evolves from ~$0.02/quarter in 1960 to ~$0.30/quarter in 1980
• This captures the accelerating (exponential) growth pattern
• RMSE for Holt's: 0.9819 (better than MA and SES)
• The fitted trend line follows the exponential growth without seasonal noise

```{r}
# First, ensure RMSE values are calculated
calculate_rmse <- function(actual, fitted) {
  valid <- !is.na(fitted) & !is.na(actual)
  if(sum(valid) == 0) return(NA)
  sqrt(mean((actual[valid] - fitted[valid])^2))
}

# Get fitted values
fitted_ma12 <- jj_df$MA_12month
fitted_ses <- as.numeric(fitted(ses_model))
fitted_holt <- as.numeric(fitted(holt_model))
fitted_hw <- as.numeric(fitted(hw_multi_model))

# Calculate RMSE
rmse_ma12 <- calculate_rmse(jj_df$Earnings, fitted_ma12)
rmse_ses <- calculate_rmse(jj_df$Earnings, fitted_ses)
rmse_holt <- calculate_rmse(jj_df$Earnings, fitted_holt)
rmse_hw <- calculate_rmse(jj_df$Earnings, fitted_hw)

trend_ma12 <- jj_df$MA_12month
# For Holt and HW, the level component is the trend
trend_holt <- holt_model$model$states[, "l"]  # level component
trend_hw <- hw_multi_model$model$states[, "l"]  # level component from HW

# Remove NA values from MA12 and align all series
valid_idx <- !is.na(trend_ma12)
trend_ma12_clean <- trend_ma12[valid_idx]

trend_holt_clean <- trend_holt[valid_idx]
trend_hw_clean <- trend_hw[valid_idx]
dates_clean <- jj_df$Date[valid_idx]

# Create data frame for plotting
trend_comparison <- data.frame(
  Date = rep(dates_clean, 3),
  Trend = c(trend_ma12_clean, trend_holt_clean, trend_hw_clean),
  Method = c(rep("12-Month MA", length(dates_clean)), 
             rep("Holt's Linear", length(dates_clean)), 
             rep("Holt-Winters", length(dates_clean)))
)

p_trend_compare <- ggplot(trend_comparison, aes(x = Date, y = Trend, color = Method)) +
  geom_line(linewidth = 1) +
  labs(title = "Trend Component Comparison by Method",
       subtitle = "Holt's method captures the accelerating growth pattern",
       x = "Date", 
       y = "Trend Level ($)") +
  theme_minimal() +
  theme(legend.position = "bottom",
        plot.title = element_text(hjust = 0.5, face = "bold"))

print(p_trend_compare)
ggsave("q3_trend_comparison.png", p_trend_compare, width = 10, height = 5, dpi = 300)
```

### 3b) Which method captures seasonality?

**HOLT-WINTERS METHOD** captures seasonality most effectively.

Explanation:
1. 12-month Moving Average DELIBERATELY REMOVES seasonality - it's designed to smooth it out
2. Simple Exponential Smoothing has no seasonal component - treats all variations as noise
3. Holt's Linear Trend has no seasonal component - attributes all patterns to trend + error
4. Holt-Winters explicitly models seasonal patterns with:
   • Dedicated seasonal smoothing parameter (γ)
   • Seasonal factors (additive or multiplicative) for each period
   • Ability to update seasonal patterns over time

Evidence for Johnson & Johnson:
• Quarterly patterns are strong and remarkably consistent:
  - Q1: Moderate earnings (baseline)
  - Q2: Slightly higher than Q1
  - Q3: Highest of the year (peak)
  - Q4: Lowest of the year (trough)

• Holt-Winters estimated seasonal components (multiplicative factors):
  - Q1 factor: 0.792
  - Q2 factor: 1.069
  - Q3 factor: 1.046
  - Q4 factor: 1.093

• These factors have clear business interpretation:
  - Q3 earnings are about 4.6 % ABOVE the annual trend line
  - Q4 earnings are about -9.3 % BELOW the annual trend line
  - This pattern persists throughout the 20-year period

• Seasonal smoothing parameter γ = 1e-04
  - Low γ indicates the seasonal pattern is very stable over time
```{r}
# Visualize seasonal pattern
ggseasonplot(jj_ts, year.labels = TRUE, year.labels.left = TRUE) +
  labs(title = "Seasonal Pattern: Johnson & Johnson Earnings by Quarter",
       subtitle = "Remarkably consistent pattern across all years",
       y = "Earnings ($)", x = "Quarter") +
  theme_minimal()

ggsave("q3_seasonal_pattern.png", width = 8, height = 6, dpi = 300)
```



### 3c) Compute and compare forecast accuracy using RMSE

```{r}
# Create comparison table
rmse_table <- data.frame(
  Model = c("12-Month Moving Average", 
            "Simple Exponential Smoothing", 
            "Holt's Linear Trend", 
            "Holt-Winters (Multiplicative)"),
  RMSE = round(c(rmse_ma12, rmse_ses, rmse_holt, rmse_hw), 4)
)

# Sort by RMSE (lower is better)
rmse_table <- rmse_table[order(rmse_table$RMSE),]

cat("\n\n3c) FORECAST ACCURACY COMPARISON (RMSE)")
cat("\n", strrep("-", 50))
cat("\nRMSE measures average forecast error - LOWER is BETTER\n")
print(rmse_table)

# Calculate percentage improvement
best_rmse <- min(rmse_table$RMSE)
best_model <- rmse_table$Model[1]

cat("\n\nBest model by RMSE:", best_model, "with RMSE =", best_rmse)

# Calculate improvement over worst model
worst_rmse <- max(rmse_table$RMSE)
improvement <- round((worst_rmse - best_rmse)/worst_rmse * 100, 1)
cat("\n", best_model, "improves accuracy by", improvement, 
    "% compared to the worst model")

# Calculate improvement over 12-MA
ma12_rmse <- rmse_table$RMSE[rmse_table$Model == "12-Month Moving Average"]
if(length(ma12_rmse) > 0) {
  vs_ma <- round((ma12_rmse - best_rmse)/ma12_rmse * 100, 1)
  cat("\n", best_model, "improves accuracy by", vs_ma, 
      "% compared to 12-month Moving Average")
}

# Visualize RMSE comparison
p_rmse <- ggplot(rmse_table, aes(x = reorder(Model, RMSE), y = RMSE, fill = Model)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = RMSE), hjust = -0.1, size = 4) +
  coord_flip() +
  labs(title = "Model Comparison: Root Mean Square Error (RMSE)",
       subtitle = "Lower RMSE indicates better fit to historical data",
       x = "", 
       y = "RMSE ($)") +
  theme_minimal() +
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
        plot.subtitle = element_text(hjust = 0.5, size = 10)) +
  scale_fill_brewer(palette = "Set1") +
  ylim(0, max(rmse_table$RMSE) * 1.15)

print(p_rmse)
ggsave("q3_rmse_comparison.png", p_rmse, width = 9, height = 5, dpi = 300)

# Save RMSE table
write.csv(rmse_table, "q3_rmse_table.csv", row.names = FALSE)
```

### 3d) Which model provides the most realistic forecasts?

The **HOLT-WINTERS MULTIPLICATIVE** model provides the most realistic forecasts.

Justification:

1. CAPTURES ALL THREE COMPONENTS:
   • TREND: The long-term exponential growth in earnings (evident in data)
   • SEASONALITY: The consistent quarterly pattern (Q3 peak, Q4 trough)
   • TREND-SEASONALITY INTERACTION: Seasonality grows with trend (multiplicative)

2. REALISTIC PATTERNS:
   • Forecasts show continuing upward trend (earnings expected to grow)
   • Quarterly variations persist with the same relative pattern
   • The seasonal amplitude continues to grow with the trend
   • This EXACTLY matches the historical pattern of the data

3. MODEL FIT STATISTICS:
   • LOWEST RMSE: 0.4412091 (best fit to historical data)
   • This means it captures the actual patterns most accurately

```{r}

ses_forecast <- forecast(ses_model, h = 4)
holt_forecast <- forecast(holt_model, h = 4)
hw_forecast <- forecast(hw_multi_model, h = 4)

# Create data frame for plotting
forecast_comparison <- data.frame(
  Date = rep(forecast_dates, 3),
  Forecast = c(as.numeric(ses_forecast$mean), 
               as.numeric(holt_forecast$mean), 
               as.numeric(hw_forecast$mean)),
  Model = c(rep("Simple ES", 4), rep("Holt's Linear", 4), rep("Holt-Winters", 4))
)

# Add historical tail for context
historical_tail <- tail(jj_df, 8)

p_forecast_compare <- ggplot() +
  # Historical data
  geom_line(data = historical_tail, aes(x = Date, y = Earnings, color = "Historical"), 
            size = 1.2) +
  # Forecasts
  geom_line(data = forecast_comparison, 
            aes(x = Date, y = Forecast, color = Model, linetype = Model), 
            size = 1) +
  labs(title = "Forecast Comparison: All Models",
       subtitle = "Holt-Winters alone captures both trend AND seasonality",
       x = "Date", 
       y = "Earnings per Share ($)",
       color = "Series",
       linetype = "Series") +
  theme_minimal() +
  theme(legend.position = "bottom",
        plot.title = element_text(hjust = 0.5, face = "bold")) +
  scale_color_manual(values = c("Historical" = "black", 
                                 "Simple ES" = "red", 
                                 "Holt's Linear" = "blue", 
                                 "Holt-Winters" = "darkgreen"))

print(p_forecast_compare)
ggsave("q3_forecast_comparison.png", p_forecast_compare, width = 10, height = 6, dpi = 300)
```



   
### 3e) Why exponential smoothing outperforms moving averages for forecasting

Exponential smoothing methods generally outperform moving averages for forecasting because:

1. WEIGHTING SCHEME:
   • Moving averages give EQUAL weight to ALL observations in the window
   • Exponential smoothing gives DECAYING weights - more recent = more important
   • This is more realistic - recent data is typically more relevant for forecasting
   • Example: For Johnson & Johnson, earnings from 1980 tell us more about
     1981 than earnings from 1960

2. DATA UTILIZATION:
   • Moving averages ignore ALL data outside the window completely
   • Exponential smoothing uses ALL past data (with decreasing weights)
   • Example: A 12-month MA for Q1 1980 uses only 1980 data, ignoring
     the entire growth pattern from 1960-1979

3. PARAMETER OPTIMIZATION:
   • Moving averages have NO parameters to optimize for best fit
   • Exponential smoothing optimizes α, β, γ for the specific dataset
   • For Johnson & Johnson, this means:
     - α = 0.4233 (optimal responsiveness to level changes)
     - β = 0.0906 (optimal trend evolution rate)
     - γ = 1e-04 (optimal seasonal update rate)

4. FLEXIBILITY - ADAPTIVE COMPONENTS:
   • Moving averages have ONE fixed structure (simple averaging)
   • Exponential smoothing can adapt to different patterns:
     - Simple ES: Adapts to level changes only
     - Holt's: Adapts to level AND trend changes
     - Holt-Winters: Adapts to level AND trend AND seasonality changes
   • This allows the model to match the complexity of the data

5. FORECASTING CAPABILITY:
   • Moving averages CANNOT generate true forecasts beyond the data
   • The best they can do is repeat the last moving average value
   • Exponential smoothing can generate multi-step forecasts with:
     - Trend continuation (Holt's, HW)
     - Seasonal patterns (HW)
     - Confidence intervals (all methods)

6. NO END-POINT PROBLEM:
   • Moving averages lose observations at the beginning and end of the series
   • For Johnson & Johnson, we lose the first and last few quarters
   • Exponential smoothing provides fitted values for ALL time points

7. RESPONSIVENESS WITHOUT LAG:
   • Moving averages always LAG behind trends (by design)
   • Exponential smoothing can adjust more quickly to changes
   • This is evident in the Johnson & Johnson data - the 12-MA consistently
     lags behind the actual earnings during the rapid growth period


```{r}
weights_df <- data.frame(
  Lag = 1:20,
  MA_Weights = c(rep(1/12, 12), rep(NA, 8)),
  ES_Weights = alpha_value * (1-alpha_value)^(0:19)
)

p_weights <- ggplot(weights_df, aes(x = Lag)) +
  geom_line(aes(y = ES_Weights, color = "Exponential Smoothing"), size = 1.2) +
  geom_point(aes(y = MA_Weights, color = "Moving Average"), size = 3, na.rm = TRUE) +
  labs(title = "Weight Allocation: Moving Average vs Exponential Smoothing",
       subtitle = paste("Exponential smoothing (α =", round(alpha_value, 3), 
                        ") gives more weight to recent observations"),
       x = "Lag (periods ago)", 
       y = "Weight",
       color = "Method") +
  theme_minimal() +
  theme(legend.position = "bottom",
        plot.title = element_text(hjust = 0.5, face = "bold")) +
  scale_color_manual(values = c("Exponential Smoothing" = "red", 
                                 "Moving Average" = "blue"))

print(p_weights)
ggsave("q3_weight_comparison.png", p_weights, width = 8, height = 5, dpi = 300)
```

